---
title: "Extra_SEMs_Established"
author: "Will Pfadenhauer"
date: "2023-09-18"
output: html_document
---
##Load packages
```{r}
require(data.table)
require(piecewiseSEM)
require(MASS)
require(DHARMa)
require(MuMIn)
require(ResourceSelection)
```

##Data Prep
```{r}
#Load data
data <- fread("#path to Oceanic_Islands_Master_Database.csv")

#Create established richness column (aka total nonnative species)
data$tot_nn_species <- data$Established_only_species + data$Invasive_species

#No need to center and scale predictor variables - pSEM does this automatically.

#Changing the units so that unstandardized estimates from the model are more useful. 
data$distance_landmass_10km <- data$Distance_to_nearest_landmass_m/10000
data$Area_hundredsqkm <- data$Area_m2/100000000
data$distance_continent_10km <- data$Distance_to_nearest_continent_m/10000
data$Latitude_of_centroid <- as.numeric(data$Latitude_of_centroid)
data$Latitude_of_centroid <- abs(data$Latitude_of_centroid)
```

#Purpose of this file
The purpose of this file is to test all possible combinations of correlated variables, to figure out which combination works best. 

There are three such pairs of covariates: maximum elevation and minimum monthly temperature (Pearson = 0.73), human population and native plant richness (Pearson = 0.80), and distance to nearest continent and distance to nearest landmass (inherently similar).

There are three pairs of correlated variables, so there are eight possible options:
000
001
010
011
100
101
110
111

For each of these combinations of variables, I'm following these steps: 
1. Fitting hypothesized pSEM
2. Checking summary for significance of existing relationships, and any new relationships that need to be added
3. Specifying the next iteration of the model based on the summary, and continuing until all relationships are significant. 
4. Plotting the path diagram. 

The R-squared and AIC values that are used to compare each of the 4 options are found within the summary as well. 

Some variable combinations will be properly fit after only 2 iterations - others take a few more tries. 

Correlated variables are commented out using #

Each pSEM consists of a series of sub-models, which are either negative binomial GLMs (glm.nb) or traditional GLMs. 

--

##OPTION 000: 

Native species
Distance to landmass
Maximum elevation
```{r}
m000 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Native_species + 
                       distance_landmass_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       #Human_population +                       
                       GDP_per_capita + 
                       Maximum_elevation_ft +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
               
              #Sub-model 2
               glm.nb(Native_species ~ 
                       Age_Ma + 
                       distance_landmass_10km +
                       Max_temp_deg_C_BIO5 +
                       Area_hundredsqkm +
                       Temp_range_deg_C_BIO7+
                       Maximum_elevation_ft +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                       data = data),              
              
              #Sub-model 3 
              #glm(Human_population ~ 
              #      Max_temp_deg_C_BIO5 + 
              #      Area_hundredsqkm +
              #      Year_of_modern_colonization,
              #    family = "poisson", data = data),
              
              #Sub-model 4
              #glm(Average_human_modification ~ 
              #      Human_population,
              #    family= gaussian, data=data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(m000, conserve = T)
#You may need to scroll to the end of this code chunk to see the summary. 

#Retain only significant variables (from the summary), and then add any new variables that are significant from tests of independence claims. 
m000_2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification +
                       Maximum_elevation_ft,
                      data = data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma,
                  data = data),
              
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
              
              
)

summary(m000_2, conserve = T)
#You may need to scroll to the end of this code chunk to see the summary. 

#Retain only significant variables (from the summary), and then add any new variables that are significant from tests of independence claims. 
m000_3 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Max_temp_deg_C_BIO5 + 
                        Average_human_modification+
                       Maximum_elevation_ft,
                      data = data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma+
                    Area_hundredsqkm,
                  data = data),
              
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
              
              
)

summary(m000_3, conserve = T)
#Everything looks good. 
#Note that I'm leaving out Max_temp ~ Average human modification since we know that the (broad scale) temperature in a place is not controlled by its human modification. 

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.61

#Draw the path diagram
plot(m000_3)

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(m000, m000_2)
```


I'm going to include fewer comments in the rest of the code chunks, but I'm following the exact same process as above. 



##OPTION 001:

Native species
Distance to landmass
Minimum temp
```{r}
m001 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Native_species + 
                       distance_landmass_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       #Human_population +                       
                       GDP_per_capita + 
                       Min_temp_deg_C_BIO6 +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
               
              #Sub-model 2
               glm.nb(Native_species ~ 
                       Age_Ma + 
                       distance_landmass_10km +
                       Max_temp_deg_C_BIO5 +
                       Area_hundredsqkm +
                       Temp_range_deg_C_BIO7+
                       Min_temp_deg_C_BIO6 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                       data = data),              
              
              #Sub-model 3 
              #glm(Human_population ~ 
              #      Max_temp_deg_C_BIO5 + 
              #      Area_hundredsqkm +
              #      Year_of_modern_colonization,
              #    family = "poisson", data = data),
              
              #Sub-model 4
              #glm(Average_human_modification ~ 
              #      Human_population,
              #    family= gaussian, data=data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(m001, conserve = T)

m001_2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                        distance_landmass_10km +
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification + 
                       Min_temp_deg_C_BIO6,
                      data = data),
            
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
              
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~ 
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
            
)

summary(m001_2, conserve = T)
#Everything looks good. 
#Note that I'm leaving out Max_temp ~ Average human modification since we know that the (broad scale) temperature in a place is not controlled by its human modification. 

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.55

#Draw the path diagram
plot(m001_2)

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(m001)
```


##OPTION 010: 
Native species
Distance to continent
Maximum elevation
```{r}
m010 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Native_species + 
                       distance_continent_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       #Human_population +                       
                       GDP_per_capita + 
                       Maximum_elevation_ft +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
               
              #Sub-model 2
               glm.nb(Native_species ~ 
                       Age_Ma + 
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5 +
                       Area_hundredsqkm +
                       Temp_range_deg_C_BIO7+
                       Maximum_elevation_ft +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                       data = data),              
              
              #Sub-model 3 
              #glm(Human_population ~ 
              #      Max_temp_deg_C_BIO5 + 
              #      Area_hundredsqkm +
              #      Year_of_modern_colonization,
              #    family = "poisson", data = data),
              
              #Sub-model 4
              #glm(Average_human_modification ~ 
              #      Human_population,
              #    family= gaussian, data=data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(m010, conserve = T)

m010_2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Average_human_modification + 
                       Maximum_elevation_ft,
                      data = data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma,
                  data = data)
              

)
summary(m010_2, conserve = T)
#Everything looks good. 

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.55

#Draw the path diagram
plot(m010_2)

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(m010)

```


##OPTION 011: 

Native species
Distance to continent
Minimum temp
```{r}
m011 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Native_species + 
                       distance_continent_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       #Human_population +                       
                       GDP_per_capita + 
                       Min_temp_deg_C_BIO6 +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
               
              #Sub-model 2
               glm.nb(Native_species ~ 
                       Age_Ma + 
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5 +
                       Area_hundredsqkm +
                       Temp_range_deg_C_BIO7+
                       Min_temp_deg_C_BIO6+
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                       data = data),              
              
              #Sub-model 3 
              #glm(Human_population ~ 
              #      Max_temp_deg_C_BIO5 + 
              #      Area_hundredsqkm +
              #      Year_of_modern_colonization,
              #    family = "poisson", data = data),
              
              #Sub-model 4
              #glm(Average_human_modification ~ 
              #      Human_population,
              #    family= gaussian, data=data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                 family = "gaussian", data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(m011)

m011_2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                      distance_continent_10km+
                       Average_human_modification + 
                       Min_temp_deg_C_BIO6+
                        Age_Ma,
                      data = data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm+
                    Latitude_of_centroid,
                 family = "gaussian", data = data)

)

summary(m011_2, conserve = T)

m011_3 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Average_human_modification + 
                       Min_temp_deg_C_BIO6,
                      data = data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm+
                    Latitude_of_centroid,
                 family = "gaussian", data = data)

)

summary(m011_3, conserve = T)

m011_4 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Average_human_modification + 
                       Min_temp_deg_C_BIO6+
                        Latitude_of_centroid,
                      data = data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm+
                    Latitude_of_centroid,
                 family = "gaussian", data = data)

)

summary(m011_4, conserve = T)
#Everything looks good. 

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.41

#Draw the path diagram
plot(m011_4)

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(m011, m011_2, m011_3)
```


##OPTION 100:

Human population
Distance to landmass
Maximum elevation
```{r}
m100 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       #Native_species + 
                       distance_landmass_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       Human_population +                       
                       GDP_per_capita + 
                       Maximum_elevation_ft +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
               
              #Sub-model 2
              # glm.nb(Native_species ~ 
              #         Age_Ma + 
              #         distance_landmass_km +
              #         Max_temp_deg_C_BIO5 +
              #         Area_hundredsqkm +
              #         Temp_range_deg_C_BIO7+
              #         Min_temp_deg_C_BIO6 +
              #         Precip_seasonality_BIO15 +
              #        Total_precip_mm_BIO12,
              #         data = data),              
              
              #Sub-model 3 
              glm(Human_population ~ 
                    Max_temp_deg_C_BIO5 + 
                    Area_hundredsqkm +
                    Year_of_modern_colonization,
                 family = "poisson", data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Human_population,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(m100, conserve = T)


m100_2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification + 
                       Maximum_elevation_ft+
                       Year_of_modern_colonization,
                      data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Max_temp_deg_C_BIO5,
                  family= gaussian, data=data),
              
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Area_hundredsqkm+
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)

)

summary(m100_2, conserve = T)
#Everything looks good. 

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.71

#Draw the path diagram
plot(m100_2)

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(m100)
```


##OPTION 101: 

Human population
Distance to landmass
Minimum temperature
```{r}
m101 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       #Native_species + 
                       distance_landmass_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       Human_population +                       
                       GDP_per_capita + 
                       Min_temp_deg_C_BIO6 +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
               
              #Sub-model 2
              # glm.nb(Native_species ~ 
              #         Age_Ma + 
              #         distance_landmass_km +
              #         Max_temp_deg_C_BIO5 +
              #         Area_hundredsqkm +
              #         Temp_range_deg_C_BIO7+
              #         Min_temp_deg_C_BIO6 +
              #         Precip_seasonality_BIO15 +
              #         Total_precip_mm_BIO12,
              #         data = data),              
              
              #Sub-model 3 
              glm(Human_population ~ 
                    Max_temp_deg_C_BIO5 + 
                    Area_hundredsqkm +
                    Year_of_modern_colonization,
                  family = "poisson", data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Human_population,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(m101, conserve = T)

m101_2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       distance_landmass_10km +
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification + 
                       Min_temp_deg_C_BIO6+
                       Year_of_modern_colonization,
                      data = data),
      
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Max_temp_deg_C_BIO5,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm+
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
              
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Area_hundredsqkm+
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
              
)

summary(m101_2, conserve = T)


m101_3 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification + 
                       Min_temp_deg_C_BIO6+
                       Year_of_modern_colonization,
                      data = data),
      
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Max_temp_deg_C_BIO5,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm+
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
              
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Area_hundredsqkm+
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
              
)

summary(m101_3, conserve = T)
#Everything looks good. 

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.58

#Draw the path diagram
plot(m101_3)

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(m101, m101_2)

```


##OPTION 110: WINNER - highest R-squared = 0.71 (tied with 100), lowest AIC (not tied)

Human population
Distance to continent
Maximum elevation
```{r}
m110 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       #Native_species + 
                       distance_continent_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       Human_population +                       
                       GDP_per_capita + 
                       Maximum_elevation_ft +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
               
              #Sub-model 2
              # glm.nb(Native_species ~ 
              #         Age_Ma + 
              #         distance_landmass_km +
              #         Max_temp_deg_C_BIO5 +
              #         Area_hundredsqkm +
              #         Temp_range_deg_C_BIO7+
              #         Min_temp_deg_C_BIO6 +
              #         Precip_seasonality_BIO15 +
              #        Total_precip_mm_BIO12,
              #         data = data),              
              
              #Sub-model 3 
              glm(Human_population ~ 
                    Max_temp_deg_C_BIO5 + 
                    Area_hundredsqkm +
                    Year_of_modern_colonization,
                 family = "poisson", data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Human_population,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                   data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(m110, conserve = T)

m110_2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification + 
                       Maximum_elevation_ft+
                       Year_of_modern_colonization,
                      data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    distance_continent_10km+
                    Max_temp_deg_C_BIO5,
                  family= gaussian, data=data),
              
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid +
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)

)

summary(m110_2, conserve = T)
#Everything looks good. 

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.71

#Draw the path diagram
plot(m110_2)

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(m110)

```


##OPTION 111: 

Human population
Distance to continent
minimum temperature
```{r}
m111 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       #Native_species + 
                       distance_continent_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       Human_population +                       
                       GDP_per_capita + 
                       Min_temp_deg_C_BIO6 +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
               
              #Sub-model 2
              # glm.nb(Native_species ~ 
              #         Age_Ma + 
              #         distance_landmass_km +
              #         Max_temp_deg_C_BIO5 +
              #         Area_hundredsqkm +
              #         Temp_range_deg_C_BIO7+
              #         Min_temp_deg_C_BIO6 +
              #         Precip_seasonality_BIO15 +
              #        Total_precip_mm_BIO12,
              #         data = data),              
              
              #Sub-model 3 
              glm(Human_population ~ 
                    Max_temp_deg_C_BIO5 + 
                    Area_hundredsqkm +
                    Year_of_modern_colonization,
                 family = "poisson", data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Human_population,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Area_hundredsqkm +
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(m111, conserve = T)


m111_2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                        distance_continent_10km+
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification + 
                       Min_temp_deg_C_BIO6+
                       Year_of_modern_colonization,
                      data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                   distance_continent_10km+
                    Max_temp_deg_C_BIO5,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
              
)

summary(m111_2, conserve = T)

m111_3 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification + 
                       Min_temp_deg_C_BIO6+
                       Year_of_modern_colonization,
                      data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                   distance_continent_10km+
                    Max_temp_deg_C_BIO5,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~ 
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
              
)

summary(m111_3, conserve = T)
#Everything looks good. 

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.58

#Draw the path diagram
plot(m111_3)

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(m111, m111_2)
```

