---
title: "Extra_SEMs_Native"
author: "Will Pfadenhauer"
date: "2023-09-18"
output: html_document
---
Load packages
```{r}
require(data.table)
require(piecewiseSEM)
require(MASS)
require(DHARMa)
require(MuMIn)
require(ResourceSelection)
```

Data Prep
```{r}
#Load data
data <- fread("[link to Oceanic_Islands_Master_Database.csv]")

#Create established richness column (aka total nonnative species)
data$tot_nn_species <- data$Established_only_species + data$Invasive_species

#No need to center and scale predictor variables - pSEM does this automatically.

#Changing the units so that unstandardized estimates from the model are more useful. 
data$distance_landmass_10km <- data$Distance_to_nearest_landmass_m/10000
data$Area_hundredsqkm <- data$Area_m2/100000000
data$distance_continent_10km <- data$Distance_to_nearest_continent_m/10000
data$Latitude_of_centroid <- as.numeric(data$Latitude_of_centroid)
data$Latitude_of_centroid <- abs(data$Latitude_of_centroid)
```

The purpose of this file is to test all possible combinations of correlated variables, to figure out which combination works best. 

There are two such pairs of covariates: maximum elevation and minimum monthly temperature (Pearson = 0.73) and distance to nearest continent and distance to nearest landmass (inherently similar).

Human population and native species richness are correlated as well, but since this model *predicts* native species richness, and human population is not one of our hypothesized covariates here, we don't have to worry about that combination. 

Two pairs of correlated variables, so there are four possible options:
00
01
10
11

For each of these combinations of variables, I'm following these steps: 
1. Fitting hypothesized pSEM
2. Checking summary for significance of existing relationships, and any new relationships that need to be added
3. Specifying the next iteration of the model based on the summary, and continuing until all relationships are significant. 
4. Plotting the path diagram. 

The R-squared and AIC values that are used to compare each of the 4 options are found within the summary as well. 

Some variable combinations will be properly fit after only 2 iterations - others take a few more tries. 

Correlated variables are commented out using #

Each pSEM consists of a series of sub-models, which are either negative binomial GLMs (glm.nb) or traditional GLMs. 

--

OPTION 00: 

Distance to nearest continent 
Maximum elevation 
```{r}
#First attempt
n00 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Area_hundredsqkm +
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Age_Ma + 
                       Total_precip_mm_BIO12 +
                       Precip_seasonality_BIO15 +
                     #  Min_temp_deg_C_BIO6 +
                       Max_temp_deg_C_BIO5 +
                       Temp_range_deg_C_BIO7,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm,
                  data = data),
              
              #Sub-model 3
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid +
                    Maximum_elevation_ft,
                  family = "gaussian", data = data),
              
              #Sub-model 4
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 5
            #  glm(Min_temp_deg_C_BIO6 ~
            #        Latitude_of_centroid,
            #      family = "gaussian",
            #      data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 7
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data)
)

summary(n00, conserve = T)
#You may have to scroll to the bottom of this code chunk to see results. 

#Retain only significant variables (from the summary), and then add any new variables that are significant from tests of independence claims. 
n00_2 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5+
                       Total_precip_mm_BIO12+
                       Latitude_of_centroid,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma,
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid +
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
              
)

summary(n00_2, conserve = T)
#Just have to add one more thing: Elevation ~ area

n00_3 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5+
                       Total_precip_mm_BIO12+
                       Latitude_of_centroid,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma +
                    Area_hundredsqkm,
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid +
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
              
)

summary(n00_3, conserve = T)
#This says add elevation~distance too. 

n00_4 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5+
                       Total_precip_mm_BIO12+
                       Latitude_of_centroid,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma +
                    Area_hundredsqkm+
                    distance_continent_10km,
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid +
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
              
)

summary(n00_4, conserve = T)
#Looks like we're done. 

#Draw the path diagram. 
plot(n00_4)

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.98

#Remove the in-progress ones, so we only have the final pSEM for this combination of variables. 
rm(n00, n00_2, n00_3)
```
This option (00) is the WINNER - highest R squared. We'll use distance to nearest continent and maximum elevation in the other file (SEMs_Native.Rmd). 


OPTION 01: 

Distance to nearest continent 
Min_temp_deg_C_BIO6

Correlated variables are commented out using #

Each pSEM consists of a series of sub-models, which are either negative binomial GLMs (glm.nb) or traditional GLMs. 
```{r}
#First attempt
n01 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Area_hundredsqkm +
                       #Maximum_elevation_ft +
                       distance_continent_10km +
                       Age_Ma + 
                       Total_precip_mm_BIO12 +
                       Precip_seasonality_BIO15 +
                       Min_temp_deg_C_BIO6 +
                       Max_temp_deg_C_BIO5 +
                       Temp_range_deg_C_BIO7,
                       data = data),              
              
              #Sub-model 2
              #glm.nb(Maximum_elevation_ft ~ 
              #      Area_hundredsqkm,
              #    data = data),
              
              #Sub-model 3
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
             #       Maximum_elevation_ft,
                  family = "gaussian", data = data),
              
              #Sub-model 4
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 7
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data)
)

summary(n01)
#You may have to scroll to the bottom of this code chunk to see results. 

#Retain only significant variables (from the summary), and then add any new variables that are significant from tests of independence claims. 
n01_2 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       distance_continent_10km +
                       Total_precip_mm_BIO12+
                       Min_temp_deg_C_BIO6 +
                       Max_temp_deg_C_BIO5+
                         Latitude_of_centroid,
                       data = data),    
              
             #Sub-model 3
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~
                   Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
)

summary(n01_2, conserve = T)

#Next iteration 
n01_3 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       distance_continent_10km +
                       Total_precip_mm_BIO12+
                       Min_temp_deg_C_BIO6 +
                       Max_temp_deg_C_BIO5+
                         Latitude_of_centroid,
                       data = data),    
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~
                   Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
)

summary(n01_3, conserve = T)
#Looks like we're done

#Draw the path diagram.
plot(n01_3)

#Check the Nagelkerke R-squared value (from the summary above) 
#R-squared = 0.94

#Remove in-progress ones, so that it only leaves the final version. 
rm(n01, n01_2)
```

I'm going to include fewer comments in the next two code chunks, but I'm following the exact same process as above. 


OPTION 10: 

Distance to nearest landmass
Maximum elevation 
```{r}
n10 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Area_hundredsqkm +
                       Maximum_elevation_ft +
                       distance_landmass_10km +
                       Age_Ma + 
                       Total_precip_mm_BIO12 +
                       Precip_seasonality_BIO15 +
                     #  Min_temp_deg_C_BIO6 +
                       Max_temp_deg_C_BIO5 +
                       Temp_range_deg_C_BIO7,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm,
                  data = data),
              
              #Sub-model 3
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid +
                    Maximum_elevation_ft,
                  family = "gaussian", data = data),
              
              #Sub-model 4
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 5
            #  glm(Min_temp_deg_C_BIO6 ~
            #        Latitude_of_centroid,
            #      family = "gaussian",
            #      data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 7
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data)
)

summary(n10, conserve = T)


n10_2 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_landmass_10km +
                       Max_temp_deg_C_BIO5 +
                       Temp_range_deg_C_BIO7+
                       Latitude_of_centroid,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma,
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid +
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 7
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid +
                    Area_hundredsqkm +
                    Maximum_elevation_ft ,
                  family = "gaussian", data = data)
)

summary(n10_2, conserve = T)

n10_3 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_landmass_10km +
                       Max_temp_deg_C_BIO5 +
                       Latitude_of_centroid,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma+
                      Area_hundredsqkm,
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid +
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
)

summary(n10_3, conserve = T)
plot(n10_3)

#R-squared = 0.97

rm(n10, n10_2)
```



OPTION 11: 

Distance to nearest landmass
Min_temp_deg_C_BIO6
```{r}
n11 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Area_hundredsqkm +
                     #  Maximum_elevation_ft +
                       distance_landmass_10km +
                       Age_Ma + 
                       Total_precip_mm_BIO12 +
                       Precip_seasonality_BIO15 +
                       Min_temp_deg_C_BIO6 +
                       Max_temp_deg_C_BIO5 +
                       Temp_range_deg_C_BIO7,
                       data = data),              
              
              #Sub-model 2
             # glm.nb(Maximum_elevation_ft ~ 
              #      Area_hundredsqkm,
              #    data = data),
              
              #Sub-model 3
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                #    Maximum_elevation_ft,
                  family = "gaussian", data = data),
              
              #Sub-model 4
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 7
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data)
)

summary(n11, conserve = T)


n11_2 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       distance_landmass_10km +
                       Age_Ma + 
                       Min_temp_deg_C_BIO6 +
                       Max_temp_deg_C_BIO5 +
                       Temp_range_deg_C_BIO7+
                       Latitude_of_centroid,
                       data = data),              
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 7
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian", data = data)
)

summary(n11_2, conserve = T)



n11_3 <-  psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       distance_landmass_10km +
                       Age_Ma + 
                       Min_temp_deg_C_BIO6 +
                       Max_temp_deg_C_BIO5 +
                       Latitude_of_centroid,
                       data = data),              
              
              #Sub-model 5
              glm(Min_temp_deg_C_BIO6 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 6
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
              
)

summary(n11_3, conserve = T)
plot(n11_3)

#R-squared = 0.93

rm(n11, n11_2)
```




