---
title: "SEMs_Established"
author: "Will Pfadenhauer"
date: "2023-09-05"
output: html_document
---

Load packages
```{r}
require(data.table)
require(piecewiseSEM)
require(MASS)
require(DHARMa)
require(MuMIn)
require(ResourceSelection)
require(ggplot2)
require(car)
require(semEff)
require(ggpubr)
require(svglite)
```

Note: some code files from this manuscript use the same names for objects. It is highly recommended that you clear your environment before running the following code (unless you're only running the plots section at the bottom). 

Data Prep
```{r}
#Load data
data <- fread("[link to Oceanic_Islands_Master_Database.csv]")

#Create established richness column (aka total nonnative species)
data$tot_nn_species <- data$Established_only_species + data$Invasive_species

#No need to center and scale predictor variables - pSEM does this automatically.

#Changing the units so that unstandardized estimates from the model are more useful. 
data$Area_hundredsqkm <- data$Area_m2/100000000
data$distance_continent_10km <- data$Distance_to_nearest_continent_m/10000
data$Latitude_of_centroid <- abs(data$Latitude_of_centroid)
```

To fit each pSEM, I'm following this process: 
1. Fitting hypothesized pSEM (see manuscript Figure S1)
2. Checking summary for significance of existing relationships, and any new relationships that need to be added
3. Specifying the next iteration of the model based on the summary, and continuing until all relationships are significant. 
4. Plotting the path diagram. 

See the file Extra_Established_SEMs.Rmd for details on how I chose between correlated variables for this file. 


Fitting the pSEMs
Each pSEM consists of a series of sub-models, which are either negative binomial GLMs (glm.nb) or traditional GLMs. 
```{r}
hyp <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       distance_continent_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       Human_population +                       
                       GDP_per_capita + 
                       Maximum_elevation_ft +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
              
              #Sub-model 3 
              glm(Human_population ~ 
                    Max_temp_deg_C_BIO5 + 
                    Area_hundredsqkm +
                    Year_of_modern_colonization,
                 family = "poisson", data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Human_population,
                  family= "gaussian", data=data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm,
                   data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)

summary(hyp, conserve = T)
#Initial check of hypotheses
```

This is attempt two - we're keeping the significant variables from hyp, and then adding in any new relationships identified by the tests of directed separation (as long as they have support from literature). 
```{r}

hyp2 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Max_temp_deg_C_BIO5 + 
                       Average_human_modification + 
                       Maximum_elevation_ft,
                      data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    distance_continent_10km+
                    Max_temp_deg_C_BIO5,
                  family= "gaussian", data=data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
)

summary(hyp2, conserve = T)
#Now that everything looks good, it's time for some bootstrapping to estimate the parameters. 

#Extract coefficients by bootstrapping (this make take a little while)
boot1 <- bootEff(hyp2, R = 10000, seed = 3, parallel = "multicore", type = "parametric")
#Note- there may be a handful of error messages about algorithms not converging. That is ok - we're running 10,000 iterations, so it's extremely likely that some of them won't work. 

effs1 <- semEff(boot1)
summary(effs1)  

#Remove direct effect of max temp on non-native species
```

Revised model based on hyp2 results
```{r}

hyp3 <- psem(
               #Sub-model 1
               glm.nb(tot_nn_species ~ 
                       Average_human_modification + 
                       Maximum_elevation_ft,
                      data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    distance_continent_10km+
                    Max_temp_deg_C_BIO5,
                  family= "gaussian", data=data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)
)

summary(hyp3, conserve = T)
#Everything still looks good. Let's try bootstrapping again. 

#Extract coefficients by bootstrapping (this make take a little while)
boot2 <- bootEff(hyp3, R = 10000, seed = 3, parallel = "multicore", type = "parametric")
#Note- there may be a handful of error messages about algorithms not converging. That is ok - we're running 10,000 iterations, so it's extremely likely that some of them won't work. 

effs2 <- semEff(boot2)
summary(effs2)  

#Everything significant

plot(hyp3)
```

Now, we're testing each component sub-model of our pSEM to make sure they are properly fit and don't violate any assumptions. 

Evaluate Sub-Model 1
```{r}
sm1 <-        glm.nb(tot_nn_species ~ 
                       Average_human_modification + 
                       Maximum_elevation_ft,
                      data = data)

options(na.action=na.fail)
aic_sm1 <- dredge(sm1)
#Lowest AIC is the one that uses all the variables. 

summary(sm1) #Note that dispersion parameter is just a tiny bit high.

#Check other model diagnostics
#Ratio of deviance explained - pretty good
(sm1$null.deviance - (deviance(sm1))) / sm1$null.deviance * 100

#Another check for overdispersion
sm1$deviance/sm1$df.residual # Fairly close to 1
sum((resid(sm1, type = "pearson"))^2) < sm1$df.residual #TRUE is good. 

#simulate residuals with DHARMa
simulationOutput_sm1 <- simulateResiduals(fittedModel = sm1, plot = T) #No significant issues with DHARMa - alright. 

plot(sm1$fitted.values, residuals.glm(sm1))
#No obvious patterns here. 
```


Evaluate Sub-Model 4
All the same steps as above, but for sub-model 4
```{r}
sm4 <-         glm(Average_human_modification ~ 
                    distance_continent_10km+
                    Max_temp_deg_C_BIO5,
                  family= "gaussian", data=data)

options(na.action=na.fail)
aic_sm4 <- dredge(sm4)

#Note: no dispersion parameters here since it's not a negative binomial glm. 

(sm4$null.deviance - (deviance(sm4))) / sm4$null.deviance * 100

simulationOutput_sm4 <- simulateResiduals(fittedModel = sm4, plot = T)

plot(sm4$fitted.values, residuals.glm(sm4))
```


Evaluating sub-model 7
```{r}
sm7 <-        glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid+
                    Area_hundredsqkm,
                  family = "gaussian",
                  data = data)

options(na.action=na.fail)
aic_sm7 <- dredge(sm7)

#Note: no dispersion parameters here since it's not a negative binomial glm. 

(sm7$null.deviance - (deviance(sm7))) / sm7$null.deviance * 100

simulationOutput_sm7 <- simulateResiduals(fittedModel = sm7, plot = T)

plot(sm7$fitted.values, residuals.glm(sm7))
```
The DHARMa plots and the residual vs fitted plot don't look good for this model. But the purpose of this project is to predict species richness, not maximum temperature, so it's very possible we're missing some predictors that would correct the DHARMa plots here. That's ok - it shouldn't significantly affect our parameters for the established species richness model. 


Plots 
Note that two of the plots here come from the SEMs_Invasive.Rmd file. 
```{r}
#These functions from StackExchange: https://stackoverflow.com/questions/59150905/is-there-a-ggplot2-analogue-to-the-avplots-function-in-r
avPlots.invis <- function(MODEL, ...) {
  
  ff <- tempfile()
  png(filename = ff)
  OUT <- car::avPlots(MODEL, ...)
  dev.off()
  unlink(ff)
  OUT }

ggAVPLOTS  <- function(MODEL, YLAB = NULL) {
  
  #Extract the information for AV plots
  AVPLOTS <- avPlots.invis(MODEL)
  K       <- length(AVPLOTS)
  
  #Create the added variable plots using ggplot
  GGPLOTS <- vector('list', K)
  for (i in 1:K) {
  DATA         <- data.frame(AVPLOTS[[i]])
  GGPLOTS[[i]] <- ggplot2::ggplot(aes_string(x = colnames(DATA)[1], 
                                             y = colnames(DATA)[2]), 
                                  data = DATA) +
                  geom_point(colour = 'black', shape=1,size=3) + 
                  geom_smooth(method = 'lm', se = TRUE, 
                              color = '#ffb703', formula = y ~ x) +
                  xlab(paste0('Predictor Residual \n (', 
                         names(DATA)[1], ' | others)')) +
                  ylab(paste0('Response Residual \n (',
                         ifelse(is.null(YLAB), 
                           paste0(names(DATA)[2], ' | others'), YLAB), ')'))+
                  theme_classic()}
  
  #Return output object
  GGPLOTS }

p1 <- ggAVPLOTS(hyp3[[1]])[1]
p1 <- p1[[1]]
p2 <- ggAVPLOTS(hyp3[[1]])[2]
p2 <- p2[[1]]

#p5 and p6 come from SEMs_Invasive.Rmd

export_fig <- ggarrange(p1, p2, p5, p6, nrow = 2, ncol = 2)
```

Export plot
```{r}
#Export the plot as an SVG file
#Need to run all of these lines at the same time. 
svglite("C:/Users/Will Pfadenhauer/OneDrive - University of Massachusetts/CHAP3_Island_Biogeo/Figures/R_exports/AV_plots.svg", width = 5, height = 5)
export_fig
dev.off()
```



