---
title: "SEMs_Native"
author: "Will Pfadenhauer"
date: "2023-09-07"
output: html_document
---

##Load packages
```{r}
require(data.table)
require(piecewiseSEM)
require(MASS)
require(DHARMa)
require(MuMIn)
require(ResourceSelection)
require(ggplot2)
require(semEff)
```

Note: some code files from this manuscript use the same names for objects. It is highly recommended that you clear your environment before running the following code. 

##Data Prep
```{r}
#Load data
data <- fread("#path to Oceanic_Islands_Master_Database.csv")

#Create established richness column (aka total nonnative species)
data$tot_nn_species <- data$Established_only_species + data$Invasive_species

#No need to center and scale predictor variables - pSEM does this automatically.

#Changing the units so that unstandardized estimates from the model are more useful. 
data$Area_hundredsqkm <- data$Area_m2/100000000
data$distance_continent_10km <- data$Distance_to_nearest_continent_m/10000
data$Latitude_of_centroid <- as.numeric(data$Latitude_of_centroid)
data$Latitude_of_centroid <- abs(data$Latitude_of_centroid)
```

To fit each pSEM, I'm following this process: 
1. Fitting hypothesized pSEM (see manuscript Figure S1)
2. Checking summary for significance of existing relationships
3. Specifying the next iteration of the model based on the summary, and continuing until all relationships are significant. 
4. Plotting the path diagram. 

See the file Extra_Native_SEMs.Rmd for details on how I chose between correlated variables for this file. 


##Fitting the pSEMs
Each pSEM consists of a series of sub-models, which are either negative binomial GLMs (glm.nb) or traditional GLMs. 
###Attempt 1
```{r}

hyp <- psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Area_hundredsqkm +
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Age_Ma + 
                       Total_precip_mm_BIO12 +
                       Precip_seasonality_BIO15 +
                       Max_temp_deg_C_BIO5 +
                       Temp_range_deg_C_BIO7,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm +
                    Age_Ma,
                  data = data),
              
              #Sub-model 3
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid +
                    Maximum_elevation_ft,
                  family = "gaussian", data = data),
              
              #Sub-model 4
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 5
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
               #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data)
)

#Extract coefficients by bootstrapping (this make take a little while)
boot1 <- bootEff(hyp, R = 10000, seed = 3, parallel = "multicore", type = "parametric")
#Note- there may be a handful of error messages about algorithms not converging. That is ok - we're running 10,000 iterations, so it's not too concerning if some of them won't work. 

effs1 <- semEff(boot1)
summary(effs1)  
```

###Attempt 2
This is attempt two - we're only keeping the significant variables from bootstrapped estimates of hyp. 
```{r}

hyp2  <- psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5,
                       data = data),              
              
              #Sub-model 2
              glm.nb(Maximum_elevation_ft ~ 
                    Age_Ma+
                    Area_hundredsqkm,
                  data = data),
              
               #Sub-model 5
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
              
)


summary(hyp2, conserve = T)

boot2 <- bootEff(hyp2, R = 10000, seed = 3, parallel = "multicore", type = "parametric")
effs2 <- semEff(boot2)
summary(effs2)  
#Everything is significant, looks good. 

summary(hyp2, conserve = T)
#All relationships are significant, and Fisher's C is insignificant. This means the model is properly fit. 

plot(hyp2)
```

##Check fit of component models
Now, we're testing each component sub-model of our pSEM to make sure they are properly fit and don't violate any assumptions. 

###Sub-Model 1
```{r}
              #Sub-model 1
sm1 <-         glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5,
                       data = data)

options(na.action=na.fail)
aic_sm1 <- dredge(sm1)
#Lowest AIC is the one that uses all the variables. 

summary(sm1) #Note that dispersion parameter is slightly high (should be close to 1).

#Check other model diagnostics
#Ratio of deviance explained - pretty good
(sm1$null.deviance - (deviance(sm1))) / sm1$null.deviance * 100

#Another check for overdispersion
sm1$deviance/sm1$df.residual # Fairly close to 1
sum((resid(sm1, type = "pearson"))^2) < sm1$df.residual #TRUE is good!

#simulate residuals with DHARMa
simulationOutput_sm1 <- simulateResiduals(fittedModel = sm1, plot = T) #No issues with DHARMa - looks really good. Issues would be shown in red.  

plot(sm1$fitted.values, residuals.glm(sm1))
#No obvious patterns here. 
```
This model does seem like it might be ever so slightly overdispersed (even after using negative binomial instead of Poisson), but the DHARMa simulations show that it still fits pretty well. See the Simulations.Rmd file for more details - we tested the effects of overdispersion on parameter estimates in this scenario, and it seems to be ok.

###Sub-Model 2
All the same steps as above, but for sub-model 2
```{r}
              #Sub-model 2
sm2 <-        glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm+
                    Age_Ma,
                  data = data)

options(na.action=na.fail)
aic_sm2 <- dredge(sm2)
#Lowest AIC is the one that uses all the variables. 

summary(sm2)  #Note that dispersion parameter looks good. 

#Proportion of deviance explained
(sm2$null.deviance - (deviance(sm2))) / sm2$null.deviance * 100

sm2$deviance/sm2$df.residual 
#Another check for overdispersion. Pretty close to 1, so looks ok. 

sum((resid(sm2, type = "pearson"))^2) < sm2$df.residual # TRUE is good. 

simulationOutput_sm2 <- simulateResiduals(fittedModel = sm2, plot = T)
#Some slight significant deviations here. Not too bad, but see note at the end of this code chunk.

plot(sm2$fitted.values, residuals.glm(sm2))
#No obvious patterns. 
```
The DHARMa plots have some small issues for this model. But the purpose of this project is to predict species richness, not elevation, so it's very possible we're missing some significant predictors of elevation that would correct the DHARMa plots here. That's ok - it shouldn't significantly affect our parameters for the native species richness model. 

###Sub-Model 5
All the same steps as above, but for sub-model 5
```{r}
sm5 <-          glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)

options(na.action=na.fail)
aic_sm5 <- dredge(sm5)
#Lowest AIC is one that uses all variables (in this case, just latitude)

#Note: no dispersion parameters here since it's not a negative binomial glm. 

#Proportion of deviance explained. 
(sm5$null.deviance - (deviance(sm5))) / sm5$null.deviance * 100 

simulationOutput_sm5 <- simulateResiduals(fittedModel = sm5, plot = T) 
#DHARMA plots - not looking great here. See note at the bottom of this code chunk. 

plot(sm5$fitted.values, residuals.glm(sm5))
#Maybe a slight downward trend? See note at bottom of this code chunk. 
```
The DHARMa plots don't look good for this model and the residual vs. fitted plot also has a slight trend. But the purpose of this project is to predict species richness, not maximum temperature, so it's very possible we're missing some predictors that would correct the DHARMa plots here. That's ok - it shouldn't significantly affect our parameters for the native species richness model. 

---
##Check to see if non-linear model works better for Sub-Model 2
Lastly, let's check to see whether a non-linear model would work better for elevation. 
Literature suggests there should be a hump-shaped trend between age and elevation, so it's possible that a GAM would be a better fit than a GLM. 
```{r}
hyp2a  <- psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5,
                       data = data),              
              
              #Sub-model 2 - Notice that this is now a GAM
             gam(Maximum_elevation_ft ~ 
                    Age_Ma+
                    Area_hundredsqkm,
                 family = negbin(1),
                  data = data),
              
               #Sub-model 5
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
              
)

hyp2_a_sum <- summary(hyp2a, conserve = T)
hyp2_sum <- summary(hyp2, conserve = T)
hyp2_a_sum$AIC #New GAM
hyp2_sum$AIC #Original GLM

#Pretty much the same. Usually, AIC differences <2 are negligible. 
```

What if we fit smoothing functions in the GAM?
```{r}
#What happens if we fit this with smoothing functions? 
#Checking to see if non-linear fit for island age works better. 
hyp2a_smo  <- psem(               
              #Sub-model 1
              glm.nb(Native_species ~ 
                       Maximum_elevation_ft +
                       distance_continent_10km +
                       Max_temp_deg_C_BIO5,
                       data = data),              
              
              #Sub-model 2
             gam(Maximum_elevation_ft ~ 
                    s(Age_Ma, bs = "cr")+
                    s(Area_hundredsqkm, bs="cr"),
                    family = negbin(1),
                  data = data),
              
               #Sub-model 5
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
              
)

hyp2_asmo_sum <- summary(hyp2a_smo, conserve = T)
hyp2_asmo_sum$AIC #New smoothed GAM
hyp2_sum$AIC #Original GLM
#This could work - this AIC is a bit lower than our original model. 
```

Increased performance can always come from overfitting, so let's check to see how the smoothed GAM looks. 
```{r}
#Let's look at it a little closer. 
gam <-     gam(Maximum_elevation_ft ~ 
                    s(Age_Ma, bs = "cr")+
                    s(Area_hundredsqkm, bs="cr"),
                    family = negbin(1),
                  data = data)     

options(na.action=na.fail)
aic_gam <- dredge(gam)
#Lowest AIC is one that uses all variables, that's good. 

#Proportion of deviance explained. 
(gam$null.deviance - (deviance(gam))) / gam$null.deviance * 100 # New smoothed gam
(sm2$null.deviance - (deviance(sm2))) / sm2$null.deviance * 100 #Original GLM: SM2 from above
#That's a good bit better - that might be where the improvement in AIC is coming from. 

#Let's check DF:
gam$df.residual #New smoothed GAM
sm2$df.residual #Original GLM
#The smoothing parameters, while explaining more variation, are using several additional DF. 

#Let's check DHARMA plots -
simulationOutput_gam <- simulateResiduals(fittedModel = gam, plot = T) #New smoothed GAM
simulationOutput_sm2 <- simulateResiduals(fittedModel = sm2, plot = T) # Original GLM
#DHARMA plots - not looking great here. See note at the bottom of this code chunk. 

#Check for patterns in residuals
plot(gam$fitted.values, residuals.glm(gam)) #New smoothed GAM
plot(sm2$fitted.values, residuals.glm(sm2)) #Original GLM
```
There are some clear issues in the DHARMA plots for the GAM that aren't present for the GLM. I think this likely means that the GAM is overfit, which is how it's achieving a much higher R-squared value from the same data. 
Some of the issues in the DHARMA plot could be solved by increasing theta from 1 to 3, but that uses several more degrees of freedom. In fact, that would use 13 degrees of freedom for a dataset that only has 46 islands, which is a gross violation of the "one-in-ten" rule. Even if this was overlooked, this doesn't solve the issue of having a clear heteroschedasticity in the residual vs fitted plot - an issue which is not shown in the residuals vs fitted plot from our original GLM. Thus, I think it makes the most sense to keep our original GLM for elevation. 
