---
title: "SEMs_Invasive"
author: "Will Pfadenhauer"
date: "2023-09-18"
output: html_document
---

##Load packages
```{r}
require(data.table)
require(piecewiseSEM)
require(MASS)
require(DHARMa)
require(MuMIn)
require(ResourceSelection)
require(ggplot2)
require(car)
require(semEff)
```

Note: some code files from this manuscript use the same names for objects. It is highly recommended that you clear your environment before running the following code (unless you're only running the plots section at the bottom). 

##Data Prep
```{r}
#Load data
data <- fread("#path to Oceanic_Islands_Master_Database.csv")

#Create established richness column (aka total nonnative species)
data$tot_nn_species <- data$Established_only_species + data$Invasive_species

#No need to center and scale predictor variables - pSEM does this automatically.

#Changing the units so that unstandardized estimates from the model are more useful. 
data$Area_hundredsqkm <- data$Area_m2/100000000
data$distance_continent_10km <- data$Distance_to_nearest_continent_m/10000
data$Latitude_of_centroid <- as.numeric(data$Latitude_of_centroid)
data$Latitude_of_centroid <- abs(data$Latitude_of_centroid)
```

To fit each pSEM, I'm following this process: 
1. Fitting hypothesized pSEM (see manuscript Figure S1)
2. Checking summary for significance of existing relationships
3. Specifying the next iteration of the model based on the summary, and continuing until all relationships are significant. 
4. Plotting the path diagram. 

See the file Extra_Invasive_SEMs.Rmd for details on how I chose between correlated variables for this file. 

##Fitting the pSEMs
Each pSEM consists of a series of sub-models, which are either negative binomial GLMs (glm.nb) or traditional GLMs. 

###Attempt 1
```{r}
hyp <- psem(
               #Sub-model 1
               glm.nb(Invasive_species ~ 
                       distance_continent_10km + 
                       Max_temp_deg_C_BIO5 + 
                       Area_hundredsqkm + 
                       Average_human_modification + 
                       Human_population +                       
                       GDP_per_capita + 
                       Maximum_elevation_ft +
                       Temp_range_deg_C_BIO7 +
                       Precip_seasonality_BIO15 +
                       Total_precip_mm_BIO12,
                      data = data),
              
              #Sub-model 3 
              glm(Human_population ~ 
                    Max_temp_deg_C_BIO5 + 
                    Area_hundredsqkm +
                    Year_of_modern_colonization,
                 family = "poisson", data = data),
              
              #Sub-model 4
              glm(Average_human_modification ~ 
                    Human_population,
                  family= gaussian, data=data),
              
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm+
                    Age_Ma,
                  data = data),
              
              #Sub-model 6
              glm(Temp_range_deg_C_BIO7 ~ 
                    Latitude_of_centroid,
                  family = "gaussian", data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 8
              glm(Precip_seasonality_BIO15 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data),
              
              #Sub-model 9
              glm(Total_precip_mm_BIO12 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)
              
#Extract coefficients by bootstrapping (this make take a little while)
boot1 <- bootEff(hyp, R = 10000, seed = 3, parallel = "multicore", type = "parametric")
#Note- there may be a handful of error messages about algorithms not converging. That is ok - we're running 10,000 iterations, so it's extremely likely that some of them won't work. 

effs1 <- semEff(boot1)
summary(effs1)  
              
```

###Attempt 2
This is attempt two - we're keeping the significant variables from hyp, but removing the rest. 
```{r}
hyp2 <- psem(
               #Sub-model 1
               glm.nb(Invasive_species ~ 
                       Max_temp_deg_C_BIO5 + 
                       Maximum_elevation_ft,
                      data = data),
               
              #Sub-model 5
              glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm+
                    Age_Ma,
                  data = data),
 
              #Sub-model 7
              glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)
)


#Extract coefficients by bootstrapping (this make take a little while)
boot2 <- bootEff(hyp2, R = 10000, seed = 3, parallel = "multicore", type = "parametric")
#Note- there may be a handful of error messages about algorithms not converging. That is ok - we're running 10,000 iterations, so it's extremely likely that some of them won't work. 

effs2 <- semEff(boot2)
summary(effs2)  
#Everything significant

summary(hyp2, conserve = T)
#Now that everything looks good, it's time for some bootstrapping to estimate the parameters. 

plot(hyp2)
```

##Check fit of component models

###Sub-Model 1
```{r}
sm1 <-        glm.nb(Invasive_species ~ 
                       Max_temp_deg_C_BIO5 + 
                       Maximum_elevation_ft,
                      data = data)

options(na.action=na.fail)
aic_sm1 <- dredge(sm1)
#Lowest AIC is the one that uses all the variables. 

summary(sm1) #Dispersion parameter is close to 1. 

#Check other model diagnostics
#Ratio of deviance explained - decent. 
(sm1$null.deviance - (deviance(sm1))) / sm1$null.deviance * 100

#Another check for overdispersion
sm1$deviance/sm1$df.residual # Nearly 1
sum((resid(sm1, type = "pearson"))^2) < sm1$df.residual #TRUE is good. 

#simulate residuals with DHARMa
simulationOutput_sm1 <- simulateResiduals(fittedModel = sm1, plot = T) #No significant issues with DHARMa - looks perfect. 

plot(sm1$fitted.values, residuals.glm(sm1))
#No obvious patterns here. 
```


###Sub-Model 5
```{r}
sm5 <-        glm.nb(Maximum_elevation_ft ~ 
                    Area_hundredsqkm+
                    Age_Ma,
                  data = data)

options(na.action=na.fail)
aic_sm5 <- dredge(sm5)
#Lowest AIC is the one that uses all the variables. 

summary(sm5) #Dispersion parameter is close to 1. 

#Check other model diagnostics
#Ratio of deviance explained - decent. 
(sm5$null.deviance - (deviance(sm5))) / sm5$null.deviance * 100

#Another check for overdispersion
sm5$deviance/sm5$df.residual # Nearly 1
sum((resid(sm5, type = "pearson"))^2) < sm5$df.residual #TRUE is good. 

#simulate residuals with DHARMa
simulationOutput_sm5 <- simulateResiduals(fittedModel = sm5, plot = T) #No significant issues with DHARMa - looks fine. Major issues would be shown in red.  

plot(sm5$fitted.values, residuals.glm(sm5))
#No obvious patterns here. 
```

###Evaluate sub-model 7
```{r}
              #Sub-model 7
sm7 <-           glm(Max_temp_deg_C_BIO5 ~
                    Latitude_of_centroid,
                  family = "gaussian",
                  data = data)

options(na.action=na.fail)
aic_sm7 <- dredge(sm7)
#Lowest AIC is the one that uses all the variables. 

#Note: no dispersion parameters here since it's not a negative binomial glm. 

#Ratio of deviance explained - decent. 
(sm7$null.deviance - (deviance(sm7))) / sm7$null.deviance * 100

#simulate residuals with DHARMa
simulationOutput_sm7 <- simulateResiduals(fittedModel = sm7, plot = T) #Not great for residuals vs. predicted, but Q-Q looks good. 

plot(sm7$fitted.values, residuals.glm(sm7)) #Seems like a possible downward trend. 
```
The DHARMa plots and the residual vs fitted plot don't look great for this model. But the purpose of this project is to predict species richness, not maximum temperature, so it's very possible we're missing some predictors that would correct the DHARMa plots here. That's ok - it shouldn't significantly affect our parameters for the invasive species richness model. 

##Create Added Variable Plots (Fig 3) 
```{r}
#These functions from StackExchange: https://stackoverflow.com/questions/59150905/is-there-a-ggplot2-analogue-to-the-avplots-function-in-r
avPlots.invis <- function(MODEL, ...) {
  
  ff <- tempfile()
  png(filename = ff)
  OUT <- car::avPlots(MODEL, ...)
  dev.off()
  unlink(ff)
  OUT }

ggAVPLOTS  <- function(MODEL, YLAB = NULL) {
  
  #Extract the information for AV plots
  AVPLOTS <- avPlots.invis(MODEL)
  K       <- length(AVPLOTS)
  
  #Create the added variable plots using ggplot
  GGPLOTS <- vector('list', K)
  for (i in 1:K) {
  DATA         <- data.frame(AVPLOTS[[i]])
  GGPLOTS[[i]] <- ggplot2::ggplot(aes_string(x = colnames(DATA)[1], 
                                             y = colnames(DATA)[2]), 
                                  data = DATA) +
                  geom_point(colour = 'black', shape=1,size=3) + 
                  geom_smooth(method = 'lm', se = TRUE, 
                              color = '#f76f73', formula = y ~ x) +
                  xlab(paste0('Predictor Residual \n (', 
                         names(DATA)[1], ' | others)')) +
                  ylab(paste0('Response Residual \n (',
                         ifelse(is.null(YLAB), 
                           paste0(names(DATA)[2], ' | others'), YLAB), ')'))+
                  theme_classic()}
  
  #Return output object
  GGPLOTS }

# p1 - p5 can be found in SEMs_Established.Rmd

p5 <- ggAVPLOTS(hyp2[[1]])[1]
p5 <- p5[[1]]
p6 <- ggAVPLOTS(hyp2[[1]])[2]
p6 <- p6[[1]]

#Save p5 and p6 so they can be opened with SEMs_Established.Rmd file
save(p5, file = "#path of where to save p5.rda")

save(p6, file = "#path of where to save p6.rda")
```
